---
title: Saving `SingleCellExperiment`s to artifacts and back again
author:
- name: Aaron Lun
  email: infinite.monkeys.with.keyboards@gmail.com
package: alabaster.sce
date: "Revised: September 22, 2022"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Saving and loading SingleCellExperiments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(BiocStyle)
self <- Githubpkg("ArtifactDB/alabaster.sce")
knitr::opts_chunk$set(error=FALSE, warning=FALSE, message=FALSE)
```

# Overview 

The `r self` package implements methods to save `SingleCellExperiment` objects to file artifacts and load them back into R.
Check out the `r Githubpkg("ArtifactDB/alabaster.base")` for more details on the motivation and concepts of the **alabaster** framework.

# Quick start

Given a `(Ranged)SingleCellExperiment`, we can use `stageObject()` to save it inside a staging directory:

```{r}
library(SingleCellExperiment)
example(SingleCellExperiment, echo=FALSE) # can't be bothered to copy it here.
sce

library(alabaster.sce)
tmp <- tempfile()
dir.create(tmp)
meta <- stageObject(sce, tmp, "sce")
.writeMetadata(meta, tmp)

list.files(tmp, recursive=TRUE)
```

We can then load it back into the session with `loadObject()`.

```{r}
meta <- acquireMetadata(tmp, "sce/experiment.json")
roundtrip <- loadObject(meta, tmp)
class(roundtrip)
```

More details on the metadata and on-disk layout are provided in the [schema](https://artifactdb.github.io/BiocObjectSchemas/html/single_cell_experiment/v1.html).

# Session information {-}

```{r}
sessionInfo()
```

